---
title: "Homework 8"
author: "Prady Kandi, EID: prk599"
date: "2025-04-14"
output:
  pdf_document:
    toc: true
urlcolor: blue
---

The link to the Github repo containing the R file can be found [here](https://github.com/PradyK756/SDS-315-HW8).

\newpage

```{r imports, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(mosaic)
library(knitr)
```

```{r setup, echo = FALSE}
creatine <- read.csv("creatinine.csv")
covid <- read.csv("covid.csv")
milk <- read.csv("milk.csv")
```

# Problem 1


## Part A

```{r, echo = FALSE}
creatinelm <- lm(creatclear ~ age, data = creatine)
coef_table <- coef(creatinelm) |> 
  round(3) |> 
  as.data.frame()


coef_table <- data.frame(
  Term = names(coef(creatinelm)),
  Estimate = round(coef(creatinelm), 3),
  row.names = NULL
)

kable(coef_table, caption = "Coefficients for Regression: Creatinine Clearance vs. Age")
```



By using a  linear regression model predicting creatine levels based off of age, the y-intercept is 147.8 and the slope is -.62. Plugging in these values for a 55 year old, we should expect a creatine level of $(147.8-55 * (.62)) = 113.7$ $mL/min$.

## Part B

Creatine clearance decreases as people get older by a rate of .62 $mL/min$ per year. This is determined by observing the slope of the regression graph.

## Part C

Using the intercept of 147.8 and slope of -.62., we can determine the expected creatine level for each age. A 40 year old has an expected creatine level of $(147.8-40(.62)) = 123.0$ $mL/min$, and a 60 year old has an expected creatine level of $(147.8-60 * (.62)) = 110.6$ $mL/min$.

Once we have these values, we can subtract the actual values to find which is higher. A 60 year old with a creatine level of 112 $mL/min$ is $112-110.6$ = 1.4 $mL/min$ higher than expected, but a 40 year old with a creatine level of 135 $mL/min$ is $135-123$ = 12 $mL/min$ higher than expected. Since $12 > 1.4$ it is concluded that the 40 year has a higher creatine level for his age and is thus healthier.

# Problem 2

```{r, echo = FALSE}
covidS <- covid |>
  filter(country  == "Spain")

covidI <- covid |>
  filter(country  == "Italy")

```

## Part 1

To find the growth rate and doubling times of each country, I first separated the dataset into each country, then applied an exponential growth model to each country's death counts. After producing the model, each dataset was resampled 10,000 times in a Monte Carlo simulation, with the doubling time and growth rate being recorded for each iteration. This allowed me to form a confidence interval for each value.\
To find the doubling time, I used the equation\
$$
\text{Doubling Time} = \frac{\log(2)}{r}
$$\
The reason I didn't use 70 as the numerator is because $r$ is already a logarithmic rate, rather than a percentage of growth.

```{r P2 Chug, echo = FALSE}
bootstrapI <- do(10000) * {
  sample_data <- resample(covidI)  
  coef(lm(log(deaths) ~ days_since_first_death, data = sample_data))  
}
growth_rate_I <- bootstrapI$days_since_first_death
doubling_time_I <- log(2) / growth_rate_I

bootstrapS <- do(10000) * {
  sample_data <- resample(covidS)  
  coef(lm(log(deaths) ~ days_since_first_death, data = sample_data))  
}
growth_rate_S <- bootstrapS$days_since_first_death
doubling_time_S <- log(2) / growth_rate_S

test1 <- 70 / 100* growth_rate_I
test2 <- 70 / 100 * growth_rate_S

```

These are the values for Italy:

Italy's Growth Rate: `r round(mean(growth_rate_I), 3)`\
Italy's Growth Rate 95% Confidence Interval: `r paste0("[", paste(round(quantile(growth_rate_I, c(0.025, 0.975)), 3), collapse = ", "), "]")`\
Italy's Doubling time: `r round(mean(doubling_time_I), 1)` days \newline Italy's Doubling time 95% Confidence Interval: `r paste0("[", paste(round(quantile(doubling_time_I, c(0.025, 0.975)), 1), collapse = ", "), "]")`

## Part 2

Like with Italy, the Spain estimates are created using a Monte Carlo simulation.

Spain's Growth Rate: `r round(mean(growth_rate_S), 3)`\
Spain's Growth Rate 95% Confidence Interval: `r paste0("[", paste(round(quantile(growth_rate_S, c(0.025, 0.975)), 3), collapse = ", "), "]")`\
Spain's Doubling time: `r round(mean(doubling_time_S), 1)` days \newline Spain's Doubling time 95% Confidence Interval: `r paste0("[", paste(format(quantile(doubling_time_S, c(0.025, 0.975)), nsmall = 1, digits = 1), collapse = ", "), "]")`


## Part 3

```{r, echo = FALSE}
covid |>
  ggplot(aes(x = days_since_first_death, y= deaths, color = country)) + labs(x = "Days Since First Death", y = "Deaths per Day", title = "Deaths per Day Over Time and per Country", color = "Country") + geom_line()
```

A graph showing the deaths per country over time.

\newpage


# Problem 3


Economists use the power-law model $Q = KP^{\beta}$, where $Q$ is quantity demanded, $K$ is a constant, $P$ is the price, and ${\beta}$ is the price elasticity of demand, which is our desired result. By taking the log of both sides, we obtain the equation

$$
\log(Q) = \log(KP^{\beta}) = \log(K) + \log(P^{\beta}) = \log(K) + \beta \log(P)
$$

This mimics a linear regression model:

$$
\log(Q) = \beta \log(P) + \log(K)
$$

where ${\beta}$ is the slope and $\log(K)$ is the constant intercept. By taking the log of price and sales and fitting a linear regression model, and then running a Monte Carlo simulation 10,000 times, we can obtain an estimate and confidence interval for the price elasticity of milk, which will be our slope.

```{r P3 Chug, echo = FALSE}
milklm <- do(10000) * coef(
  lm(log(sales) ~ log(price), data = mosaic::resample(milk))
)
```


```{r, echo = FALSE}
slopeCI <- quantile(milklm$log.price., probs = c(0.025, 0.975))
```


Milk Price Elasticity Estimate ($\beta$): `r round(mean(milklm$log.price.), 3)`  
95% Confidence Interval: `r paste0("[", paste(round(quantile(milklm$log.price., c(0.025, 0.975)), 3), collapse = ", "), "]")`
